Incorporating Periodic Variability in Hidden Markov Models for Animal Movement

### 2016-08-02 bmc_template/figures.Rnw
just need figure captions and then done

<<r msdlist>>=
get_cat_pars2 <- function(catid){
  templist <- readRDS(paste("../summary_stats/cat.",catid,".df.RDS",sep=""))
  moddf <- templist$msdlist

@@

all_cat_pars2 <- ldply(lme4:::namedList(1,22,14,15),get_cat_pars2)

@@

allmodsumdf2 <- readRDS("../summary_stats/cat.14.df2.RDS")$multimoddf
df2 <- transform(allmodsumdf2, #change allmodsumdf to allmodsumdf2
            distrib=ifelse(grepl("W",model),"Weibull","log-Normal"),
            turn_angle=ifelse(grepl("VM",as.character(model)),"yes","no"),
            homog=ifelse(grepl("TH",type),"time-geneous HMM","homogeneous HMM"))
(ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))
    + geom_line()
    + facet_wrap(~ homog)
    + scale_colour_brewer(name="step length\ndistribution",palette="Set1")
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    + scale_linetype(name="von Mises\nturning angle?")
    + labs(y=expression(Delta*"BIC"),
         x="Number of States")
    + geom_hline(yintercept=0,colour="gray")
    + zmargin)
@

@@

linetypev <- c(rep(2,2),rep(1,5))
sumdat <- transform(readRDS("../summary_stats/cat.14.df2.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

@@

avgdf <- readRDS("../summary_stats/cat.14.df2.RDS")$avgplotdf
obsdf <- (readRDS("../summary_stats/cat.14.RDS") 
  %>% dplyr::select(c(LogDist,Time)) 
  %>% filter(!is.na(LogDist)) 
  %>% group_by(Time) 
  %>% summarise_each(funs(mean))
  %>% transmute(time = Time, variable = "obs", value=LogDist)
)
vit <- readRDS("../summary_stats/cat.14.vit.RDS")
vitdf <- (vit %>% dplyr::select(c(viterbi,time))
  %>% group_by(time)
  %>% summarise_each(funs(mean))
  %>% mutate(model="homogeneous (n=3)")
)

@@

@@

>>>>

<<r msdlist,message=FALSE, echo=FALSE, warning=FALSE>>=
get_cat_pars2 <- function(catid){
  templist <- readRDS(paste("../summary_stats/cat.",catid,".df.RDS",sep=""))
  moddf <- templist$msdlist

@@

all_cat_pars2 <- ldply(lme4:::namedList(1,2,14,15),get_cat_pars2)

@@

allmodsumdf2 <- readRDS("../summary_stats/cat.14.df.RDS")$multimoddf
df2 <- transform(allmodsumdf2, #change allmodsumdf to allmodsumdf2
            distrib=ifelse(grepl("W",model),"Weibull","log-Normal"),
            turn_angle=ifelse(grepl("VM",as.character(model)),"yes","no"),
            homog=ifelse(grepl("TH",type),"time-geneous HMM","homogeneous HMM"))
(ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))
    + geom_line()
    + facet_wrap(~ homog)
    + scale_colour_brewer(name="step length\ndistribution",palette="Set1")
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    + scale_linetype(name="von Mises\nturning angle?")
    + labs(y=expression(Delta*"BIC"),
         x="Number of States")
    + geom_hline(yintercept=0,colour="gray")
    + zmargin)
@

@@

linetypev <- c(rep(2,2),rep(1,5))
sumdat <- transform(readRDS("../summary_stats/cat.14.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

@@

avgdf <- readRDS("../summary_stats/cat.14.df.RDS")$avgplotdf
obsdf <- (readRDS("../summary_stats/cat.14.RDS") 
  %>% dplyr::select(c(LogDist,Time)) 
  %>% filter(!is.na(LogDist)) 
  %>% group_by(Time) 
  %>% summarise_each(funs(mean))
  %>% transmute(time = Time, variable = "obs", value=LogDist)
)
vit <- readRDS("../summary_stats/cat.14.vit.RDS")
vitdf <- (vit %>% dplyr::select(c(viterbi,time))
  %>% group_by(time)
  %>% summarise_each(funs(mean))
  %>% mutate(model="homogeneous (n=3)")
)

@@

mnames <- c("fmm4","fmmsin3","hmm5","hmmblock4","hmmhourly3",
            "hmmquad5","hmmsin4","obs","viterbi")
mnames2 <- c("fmm-homogeneous  (n=4)"
             , "fmm-sin (n=3)"
             , "homogeneous (n=5)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=5)"
             , "sin (n=4)"
             , "observed"
             , "viterbi homogenous (n=3)")

@@

\clearpage
%cat15

<<BICred_plot15,warning=FALSE,echo=FALSE,message=FALSE,fig.width=6,fig.height=4>>=

allmodsumdf2 <- readRDS("../summary_stats/cat.15.df.RDS")$multimoddf
df2 <- transform(allmodsumdf2, #change allmodsumdf to allmodsumdf2
            distrib=ifelse(grepl("W",model),"Weibull","log-Normal"),
            turn_angle=ifelse(grepl("VM",as.character(model)),"yes","no"),
            homog=ifelse(grepl("TH",type),"time-geneous HMM","homogeneous HMM"))
(ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))
    + geom_line()
    + facet_wrap(~ homog)
    + scale_colour_brewer(name="step length\ndistribution",palette="Set1")
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    + scale_linetype(name="von Mises\nturning angle?")
    + labs(y=expression(Delta*"BIC"),
         x="Number of States")
    + geom_hline(yintercept=0,colour="gray")
    + zmargin)
@

<<adj_BIC_comparisons15, warning=FALSE, echo=FALSE, message=FALSE,,fig.width=8>>=
#L <- load("cat1sumdat.RData")

linetypev <- c(rep(2,2),rep(1,5))
sumdat <- transform(readRDS("../summary_stats/cat.15.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

<<avg_step_length_by_time15, message=FALSE, echo=FALSE, warning=FALSE,fig.width=8>>=

avgdf <- readRDS("../summary_stats/cat.15.df.RDS")$avgplotdf
obsdf <- (readRDS("../summary_stats/cat.15.RDS") 
  %>% dplyr::select(c(LogDist,Time)) 
  %>% filter(!is.na(LogDist)) 
  %>% group_by(Time) 
  %>% summarise_each(funs(mean))
  %>% transmute(time = Time, variable = "obs", value=LogDist)
)
vit <- readRDS("../summary_stats/cat.15.vit.RDS")
vitdf <- (vit %>% dplyr::select(c(viterbi,time))
  %>% group_by(time)
  %>% summarise_each(funs(mean))
  %>% mutate(model="homogeneous (n=3)")
)

mnames <- c("fmm3","fmmsin4","hmm6","hmmblock4","hmmhourly3",
            "hmmquad5","hmmsin5","obs","viterbi")
mnames2 <- c("fmm-homogeneous  (n=3)"
             , "fmm-sin (n=4)"
             , "homogeneous (n=5)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=5)"
             , "sin (n=5)"
             , "observed"
             , "viterbi homogenous (n=3)")

<<acf_plot15, message=FALSE, echo=FALSE, warning=FALSE>>=
acffun <- function(sim){
  a <- acf(sim,na.action=na.pass,plot=FALSE)
  df <- data.frame(ACF=c(a$acf),lag=a$lag)
  return(df)
}
vitacf <- (vit %>% dplyr::select(-c(time,obs))%>%ldply(.,acffun) %>% mutate(model=.id)
)

acfdf <- (readRDS("../summary_stats/cat.15.df.RDS")$acfdf 
          %>% mutate(model = .id)
)
